\mythesistitleNOTSYNPOSIS

### 2012-01-12 src/010chap1.tex
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

>>>>

\begin{definition}
  A {\em permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
  sometimes, $\lambda$ is written as a sequence $x_{i_1}x_{i_2}\ldots
  x_{i_n}$, where $i_j \in \{1, 2 \ldots n\}$ to mean that
  $\lambda(x_j)= x_{i_j}$.\\
  This document uses both notations as convenient in context.
\end{definition}

### 2012-01-12 src/020chap2.tex
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

\begin{definition}%
  \label{def:matrixsetsystem}%
  % [\emph{Set system of a binary matrix}.]
  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
    matrix} $M$.\\
  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
  \cop if it is possible to assign a linear order $\lambda$ to $U$
  where each set $S \in \cF$ appears as a consecutive subsequence of
  $\lambda$ (an interval) in the linear order.
\end{definition}

@@

Generalized \PQtree is a data structure defined in \cite{n89, mcc04} for
arbitrary set systems. If the set system has \COP, the generalized
\PQtree is identical to its \PQtree. We will use the notation from
\cite{n89} and call generalized \PQtrees as \gPQtrees.

A \gPQtree of a given set system $\cF$ from universe $U$ represents
all the sets from $\power{U}$ that have trivial intersections with
every set in $\cF$.

Two sets $A$ and $B$ are said to have a trivial intersection if $A
\cap B$ is one of the following -- $\emptyset$, $A$ or $B$. In other
words, $A$ and $B$ are either disjoint or one is the subset of the
other. {\em Trivial sets} are sets that have trivial intersections
with any set in $\power{U}$. These are namely, $U$, singleton sets in
$\power{U}$ and $\emptyset$\cite{n89, mm96}.

An important observation made by \cite{mm96} is that if $\cF$ is a set
system with \COP then, aside from every set in $\cF$ being consecutive
(after applying the \COP order) the following sets must also be
consecutive for any $A, B \in \cF$.
\begin{enumerate}
  \item The intersection $A \cap B$ 
  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
\end{enumerate}

Additionally the {\em trivial sets} of universe $U$ denoted by
$\cT(U)$, are namely -- $U$, singleton sets and the empty set
$\emptyset$. Clearly, $\cT(U)$ are consecutive
in all orderings of $U$. $\emptyset$ is considered consecutive by
convention.

@@

For any set system $\cF$, the smallest super set system containing
$\cF$ and the trivial sets that is closed under the above operations
is called the {\em weak closure} of $\cF$, denoted $\cW(\cF)$
\cite[Def.~3.2]{mcc04} \cite[Def.~2]{mm96}.

What is interesting is that the set of \COP permutations of $\cF$ is
the same as that of $\cW(\cF)$. \cite[Th.~3]{mm96}

>>>>

\begin{definition}%
  \label{def:matrixsetsystem}%
  % [\emph{Set system of a binary matrix}.]
  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
    matrix} $M$.\\
  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
  {\em \cop} if it is possible to assign a linear order, \ie a
  permutation $\sigma = u_1u_2\ldots u_n$ to $U$ \stt each set $S \in
  \cF$ appears as a consecutive subsequence $u_{i}u_{i+1}\ldots
  u_{i+k-1}$ (also called an interval) of $\lambda$ in the linear order, where
  positive integers $k = |S|$ and $i \le n+1-k$.
\end{definition}

@@

Generalized \PQtree or \gPQtree is a data structure defined in \cite{n89} to
represent all orthogonal sets of a set system $\cF$. {\em Orthogonal
  sets} of $\cF$ are sets that do not overlap with any set in $\cF$. A
data structure with the same name was later defined in \cite{mcc04} as
part of a {\em substitution decomposition} for a set system
$\cF$. \tnote{define substitutio decomp.} This is later used to define
a certificate if $\cF$ has no \COP. \PQRtree is a data structure
defined in \cite{mm96} as a data structure to represent any set system
$\cF$ with additional information in their $R$-nodes if $\cF$ has no
\COP.  All three of these data structures are proposed as
generalizations of \cite{bl76}'s \PQtree and hence produce the
$\PQtree$ if $\cF$ has \COP. As a whole, all these three data
strutures are near-identical with their differences being notional.
In this section, we will discuss the basic theory that they all hold
and stick to the name of the data structure as \PQRtree.

A \PQRtree of a given set system $\cF$ from universe $U$ represents
all the sets from $\power{U}$ that have trivial intersections with
every set in $\cF$.  Two sets $A$ and $B$ are said to have a {\em
  trivial intersection} if $A \cap B$ is one of the following --
$\emptyset$, $A$ or $B$. In other words, $A$ and $B$ are either (i)
disjoint or (ii) one is the subset of the other. Such sets are also
called {\em orthogonal sets} or said that $A$ is orthogonal to $B$. {\em Trivial subsets}
of a universe $U$, denoted by $\cT(U)$, are sets that have trivial
intersections with any set in $\power{U}$. Namely, $\cT(U)$ consists
of $U$, singleton sets in $\power{U}$ and $\emptyset$\cite{n89, mm96}.
In other words, the all sets presented in a \PQRtree share the
property that they do not overlap with any set in $\cF$. We will now
see how this helps in testing for \COP in $\cF$.

An important observation made by \cite{mm96} is that if $\cF$ is a set
system with \COP then, aside from every set in $\cF$ being consecutive
(after applying the \COP order) the following sets must also be
consecutive for any $A, B \in \cF$.
\begin{enumerate}
  \item The intersection $A \cap B$ 
  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
\end{enumerate}
Also note that trivially, sets in $\cT(U)$ are consecutive in any
permutation of $U$\endnote{$\emptyset$ is considered consecutive by
  convention.}.

Any set system that is closed under the above operations and contains
$\cT(U)$ is called {\complete} \cite{mm96} or a {\em weakly partative
  family} \cite{mcc04}. Such a set system is easier to
work with for reasons that will become apparent.\tnote{VAGUe!}
For any set system $\cF$, the smallest super set system that is
complete/weakly partative is called the {\em weak closure} of $\cF$,
denoted $\cW(\cF)$ \cite[Def.~3.2]{mcc04}. \cite[Def.~2, smallest {\em
  complete} supercollection]{mm96}.

What is interesting is that the set of \COP permutations of $\cF$ is
the same as that of $\cW(\cF)$ \cite[Th.~3]{mm96}. The reason this is
interesting is because it becomes apparent by this theory that it is
easier to use $\cW(\cF)$ to create the $\PQRtree$ for $\cF$.

@@
